---
title: "Auto-annotation"
author: "mbrauer"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output") })
output:
  BiocStyle::pdf_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Auto-Annotate

Running this script will annotate a set of genes and variants from the Maze catalog.
Each section will generate a set of artifacts and publish these to CKAN.


Annotations will comprise:

1. **Gene** Tissue-specific expression (RNA and protein). Sources: GTEx and HPA.
2. **Gene** Cell-type-specific expression for a tissue of interest. Sources: KPMP, Broad Single Cell Portal, HPA.
3. **Variant** PheWAS readouts, general. Sources: UKB, Finngen, G&H, OpenTargets.
4. **Variant** PheWAS readouts, tissue-focused. Sources: UKB, Finngen, G&H, OpenTargets.
5. **Variant** MAF in various populations. Sources: UKB, Finngen, G&H, OpenTargets.
6. **Gene and Variant** Mendelian phenotypes. Sources: OMIM, ClinGen, ClinVar, OpenTargets.
7. **Gene** Mouse phenotype variants. Sources: Monarch, KOMP, OpenTargets
8. **Gene** Molecular function of gene. Sources: Uniprot, OpenTargets
9. **Variant** eQTL effect on RNA levels in specific tissues. Sources: Szustak lab kidney atlas, Neptune
10. **Gene** Pathway.
11. **Gene** Tool compounds.

Each of the output will generate one or more resources in the "Gene Summaries"
or "Variant Summaries" package within the Sightline CKAN organization.


```{r setup, echo=FALSE, include=TRUE, eval=FALSE, message=FALSE}
knitr::opts_chunk$set(eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE,
                      fig.width=12, fig.height=8)

library(rlang)
library(tidyverse)
library(janitor)
library(ckanr)
library(jsonlite)
library(TissueEnrich) # look at Gordon Smyth's reimplemtation of GOseq
library(kableExtra)
library(flextable)
library(edgeR)
library(AnnotationHub)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(org.Hs.eg.db)
library(rtracklayer)
library(trackViewer)
library(VariantEngine)
devtools::load_all() # to be replaced with library(AnnotationEngine)

data_dir <- "/data/AnnotationEngine"

out_dir <- tempdir()

# Set S3 staging directory and AWS profile: requires a profile of `profile_name` in `~/.aws/config` file
s3_staging_dir <- 's3://aws-athena-query-results-169945233738-us-west-2/'
user_profile <- 'superuser'

# suppress breaking error
options(rstudio.connectionObserver.errorsSuppressed = TRUE)

# Set CKAN session
ckan_url <- Sys.getenv("CKAN_URL")
ckan_key <- Sys.getenv("CKAN_KEY")
ckanr::ckanr_setup(url = ckan_url, key = ckan_key)

ah <- AnnotationHub()
edb_Hs <- ah[["AH109606"]]
edb_Mm <- ah[["AH109655"]]
txdb_Hs <- ah[["AH111585"]]
txdb_Mm <- ah[["AH75762"]]
org_Hs <- ah[["AH111575"]]
```




```{sh get-credentials, eval=FALSE, echo=FALSE}
aws sso login
```


```{r user-input}
filter_context <- "Renal New Targets 2023" # "all"
filter_tissue <- "kidney" # "all"
```

## Get genes and variants from CKAN catalog

Gene and variant lists, and summary data about these, are in the "Sightline"
organization of CKAN. The organization includes packages for the Gene Catalog
(GC) and the Variant Catalog (VC), with resources for

* gene and variant lists
* summaries over gene sets and variant sets
* individual gene summaries

```{r sightline-organization}
# CKAN organizations

sightline_org_id <- "acd65bfd-df6d-4297-ae33-ab837864b022"
sightline_catalog_pkg_id <- "6035361d-871c-4e24-a027-e044fe5d42a7"
sightline_genes_catalog_res_id <- "7f7d302e-f5ab-49e2-812d-8d77473e9ab0"
sightline_internal_variants_catalog_res_id <- "fbddef6d-b3b3-4386-933d-50be682648e6"
sightline_variants_catalog_res_id <- "b79b4ae1-cdd4-499b-a9b5-2ade04a4b8e2"

new_targets_org_id <- "6b312964-c98e-40ce-a8be-e9516c4f057d"
new_targets_org <- prepare_for_publication(ckan_org_id = new_targets_org_id,
                                    url = ckan_url,
                                    key = ckan_key,
                                    dry_run = TRUE)



test_org_id <- "d4c31ff0-d146-456f-8114-f19e44bf8dbf"
test_org <- prepare_for_publication(ckan_org_id = test_org_id,
                                    url = ckan_url,
                                    key = ckan_key,
                                    dry_run = FALSE)
```


We start by loading targets and variants from the CKAN datastore. This results
in simple lists of identifiers, with minimal annotation. In the case of genes,
the gene symbol and the gene's "context" (i.e., which program it is relevant to)
are shown. For variants the associated gene_id and symbol are given.


Fetching variations associated to the selected genes from the variant catalog on CKAN.

```{r loading-and-initial-annotation}
## 0. Initial annotation
genelist <- ckanr::ds_search(resource_id = sightline_genes_catalog_res_id,
                          as="table")$records %>%
    tibble::as_tibble() %>%
    dplyr::select(-`_id`) %>%
    dplyr::filter(if(filter_context != "all") context == filter_context else TRUE) %>%
    distinct()

gene_annotation <- annotate_genes(genelist,
                                  setname = filter_context,
                                  tissue = filter_tissue,
                                  data_dir = data_dir,
                                  out_dir = out_dir,
                                  edb = edb_Hs)

variantlist <- ckanr::ds_search(resource_id = sightline_variants_catalog_res_id,
                                limit = 25000,
                                as="table")$records %>%
  tibble::as_tibble() %>%
  dplyr::select(-`_id`)

variant_annotation <- annotate_variants(genelist %>%
                                            left_join(variantlist %>%
                                                          dplyr::select(-symbol), by="gene_id") %>%
                                            dplyr::select(label, gene_id, symbol, everything()) %>%
                                            dplyr::filter(!is.na(label)) %>%
                                            dplyr::rename(variant_id = label),
                                        setname = filter_context,
                                        tissue = filter_tissue,
                                        data_dir = data_dir,
                                        edb = edb_Hs)

```

## 1. Tissue-specific expression (RNA and protein). Sources: GTEx and HPA.

Output 1:
  * GTEx: Box plot of normalized expression (RPKM) across tissues for single gene [could this be a table?]
  * HPA: table of protein expression score across all tissues
Output 2:
  * GTEx: Heatmap of normalized expression (RPKM) of all genes from genetics scan across all tissues (expression shown as relative to median across tissues for each gene)
  * HPA: Heatmap of protein expression score of all genes from genetics scan across all tissues (HPA)

```{r tissue-expression}
## 1a. Tissue specificity
gene_annotation <- annotate_gene_tissue_rna_specificity(gene_annotation)
gene_annotation <- publish_annotation(gene_annotation, ckan_org = new_targets_org, prefix = "01a")
```

```{r tissue-expression}
## 1b. Tissue specificity
gene_annotation <- annotate_gene_tissue_hpa_specificity(gene_annotation)
gene_annotation <- publish_annotation(gene_annotation, ckan_org = new_targets_org, prefix = "01b")
```

```{r cell-type-expression}
## 2. Cell type specificity within tissue
gene_annotation <- annotate_gene_cell_type_specificity_kpmp(gene_annotation)
gene_annotation <- publish_annotation(gene_annotation, ckan_org = new_targets_org, prefix = "02")
```

```{r phewas-general}
## 3. PheWAS readouts, general
variant_annotation <- annotate_variant_phewas_general(variant_annotation)
variant_annotation <- publish_annotation(variant_annotation, ckan_org = new_targets_org, prefix = "03")
```

```{r phewas-focused}
## 4. PheWAS readouts, tissue-focused
variant_annotation <- annotate_variant_phewas_focused(variant_annotation)
variant_annotation <- publish_annotation(variant_annotation, ckan_org = new_targets_org, prefix = "04")
```

```{r maf}
## 5. MAF in various populations
variant_annotation <- annotate_variant_maf(variant_annotation)
variant_annotation <- publish_annotation(variant_annotation, ckan_org = new_targets_org, prefix = "05")
```

```{r gene-mendelian-phenotypes}
## 6a. Mendelian phenotypes (gene)
gene_annotation <- annotate_gene_phenotypes(gene_annotation)
gene_annotation <- publish_annotation(gene_annotation, ckan_org = new_targets_org, prefix = "06a")
```

```{r variant-mendelian-phenotypes}
## 6b. Mendelian phenotypes (variant)
variant_annotation <- annotate_variant_phenotypes(variant_annotation)
variant_annotation <- publish_annotation(variant_annotation, ckan_org = new_targets_org, prefix = "06b")
```

```{r mouse-phenotypes}
## 7. Mouse phenotypes
gene_annotation <- annotate_mouse_gene_phenotypes(gene_annotation)
gene_annotation <- publish_annotation(gene_annotation, ckan_org = new_targets_org, prefix = "07")
```

```{r molecular-function}
## 8. Molecular function
gene_annotation <- annotate_gene_molecular_function(gene_annotation)
gene_annotation <- publish_annotation(gene_annotation, ckan_org = new_targets_org, prefix = "08")
```

```{r pathways}
## 10. Pathways
gene_annotation <- annotate_gene_pathways(gene_annotation)
gene_annotation <- publish_annotation(gene_annotation, ckan_org = new_targets_org, prefix = "10")
```

```{r tool-compounds}
## 11. Tool compounds
gene_annotation <- annotate_gene_tool_compounds(gene_annotation)
gene_annotation <- publish_annotation(gene_annotation, ckan_org = new_targets_org, prefix = "11")
```

```{r clean-up}

# close connection

```


```{r load-expression-resources-from-local-cache, eval = FALSE}

# this is for development only--these resources to come from the data lake



```

### Output 1: GTEx box plot of normalized expression (RPKM) across tissues for each gene

Code generates plots and saves to CKAN in Sightline organization, "Gene Summaries"
package. Each plot is a resource.

```{r gtex-box-plots}
tissue_expression_pkg_id <- "c86a22a7-b293-4c56-ab8a-1e9620a2b388"

symbollist <- genes %>%
  filter(context == filter_context) %>%
  distinct(symbol) %>%
  pull(symbol) 

rpkms <- gtex_v8_rpkms %>%
  dplyr::filter(gene_id %in% (genes %>%
                  filter(context == filter_context) %>%
                  distinct(gene_id) %>%
                  pull(gene_id))) %>%
  pivot_longer(cols = c(-gene_id, -gene_id_version, -symbol), names_to = "SAMPID", values_to = "expression") %>%
  left_join(gtex_v8_samples %>% dplyr::select(SAMPID, SMTS, SMTSD), by="SAMPID") %>%
  dplyr::rename(Tissue = SMTS, Subtissue = SMTSD)

lapply(symbollist, function(gene) {
    tempfile <- fs::path(tempdir, paste0(gene, "_GTEx_tissue_expn_boxplot.png"))
    p <- rpkms %>%
        dplyr::filter(symbol == gene) %>%
        ggplot(aes(x=Tissue, y=expression, group=Tissue)) +
        geom_boxplot() +
        ggtitle(gene) +
        ylab("GTEx RPKMs") +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
    # png(tempfile)
    print(p)
    dev.off()
    # ckanr::resource_create(package_id = tissue_expression_pkg_id,
    #                      name = paste0(gene, "_GTEx_tissue_expn_boxplot"),
    #                      description = paste0("GTEx tissue expression for", gene),
    #                      upload = tempfile,
    #                      url = ckan_url,
    #                      key = ckan_key)
    # unlink(tempfile)
})
```

### Output 1: plot and table of HPA protein expression score across all tissues

Generates and publishes a table of tissue-specific protein expression.

```{r tissue-protein-expression-heatmap}

p <- genes %>%
  dplyr::filter(context == filter_context) %>%
  left_join(normal) %>%
  dplyr::select(symbol, Tissue, cell_type, value) %>%
  readr::write_csv(fs::path(tempdir, "tissue_protein_expression.csv")) %>%
  ggplot(aes(symbol, Tissue)) + geom_tile(aes(fill = value),
                                              colour = "white") + scale_fill_gradient(low = "white",
                                                                                      high = "steelblue")+
  labs(x='', y = '') +
  theme_bw() +
  ggtitle(paste("Tissue protein detection for genes from", filter_context)) +
  guides(fill = guide_legend(title = "Protein level")) +
  #theme(legend.position="none") +
  theme(plot.title = element_text(hjust = 0.5,size = 20),axis.title = element_text(size=15))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),panel.grid.major= element_blank(),panel.grid.minor = element_blank())

tempfile <- fs::path(tempdir, "tissue_protein_expn_renal_new_targets.png")
png(tempfile, width = 960)
print(p)
while (!is.null(dev.list()))  dev.off()
ckanr::resource_create(package_id = tissue_expression_pkg_id,
                     name = "renal_new_targets_tissue_protein_expn_heatmap",
                     description = paste("Protein detection by tissue for genes",
                                         "in the Renal New Targets 2023 campaign (HPA)"),
                     upload = tempfile,
                     url = ckan_url,
                     key = ckan_key)
unlink(tempfile)


```

```{r tissue-gene-expression-heatmap}

p <- genes %>%
  dplyr::filter(context == filter_context) %>%
  left_join(consensus) %>%
    
tpm %>%
  dplyr::select(symbol, tissue, value) %>%
  dplyr::mutate(logTPM = log(value + 0.0001)) %>%
  readr::write_csv(fs::path(tempdir, "tissue_gene_expression.csv")) %>%
  ggplot(aes(symbol, tissue)) + geom_tile(aes(fill = logTPM),
                                              colour = "white") + scale_fill_gradient(low = "white",
                                                                                      high = "steelblue")+
  labs(x='', y = '') +
  theme_bw() +
  ggtitle(paste("Tissue expression of genes from", filter_context)) +
  guides(fill = guide_legend(title = "Log(nTPM)")) +
  #theme(legend.position="none") +
  theme(plot.title = element_text(hjust = 0.5,size = 20),axis.title = element_text(size=15))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),panel.grid.major= element_blank(),panel.grid.minor = element_blank())


tempfile <- fs::path(tempdir, "tissue_gene_expn_renal_new_targets.png")
png(tempfile, width = 960)
print(p)
while (!is.null(dev.list()))  dev.off()
ckanr::resource_create(package_id = tissue_expression_pkg_id,
                     name = "renal_new_targets_tissue_gene_expn_heatmap",
                     description = paste("Gene expression by tissue for genes",
                                         "in the Renal New Targets 2023 campaign (HPA)"),
                     upload = tempfile,
                     url = ckan_url,
                     key = ckan_key)
unlink(tempfile)

```

## 2.

```{r get-kpmp-data-from-local-cache}

cell_type_levels <- kpmp_pseudobulk %>%
    dplyr::mutate(structure = factor(tolower(structure),
                                        levels = c("vessels",
                                                   "interstitium",
                                                   "proximal tubules",
                                                   "intermediate tubules",
                                                   "distal tubules",
                                                   "collecting tubules",
                                                   "renal corpuscle"))) %>%
    dplyr::group_by(structure) %>%
    dplyr::mutate(structure_cell_type = paste(structure, cell_abbrev, sep="_")) %>%
    arrange(cell_abbrev, .by_group=TRUE) %>%
    ungroup() %>%
    dplyr::select(structure_cell_type) %>%
    distinct() %>%
    tidyr::separate_wider_delim(structure_cell_type, delim="_", names = c("structure", "cell_type")) %>%
    pull(cell_type)

gene <- "SLC6A19"
kpmp_pseudobulk %>%
    dplyr::filter(symbol == gene) %>%
    dplyr::mutate(structure = factor(tolower(structure),
                                        levels = c("vessels",
                                                   "interstitium",
                                                   "proximal tubules",
                                                   "intermediate tubules",
                                                   "distal tubules",
                                                   "collecting tubules",
                                                   "renal corpuscle")),
                  cell_type = factor(cell_abbrev,
                                    levels = cell_type_levels)) %>%
    ggplot(aes(x = cell_type, y = cpm, fill = structure)) +
    geom_boxplot() +
    ggtitle(paste("KPMP CPMs,", gene)) +
    xlab("Cell type") +
    ylab("CPMs") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

### Within specific tissue

```{r tissue-specific-gene-expression, eval=FALSE}

# data to come from KPMP and other sources

genes %>%
  dplyr::filter(context == filter_context) %>%
  left_join(consensus) %>%
  dplyr::filter(Tissue == filter_tissue) %>%
  dplyr::select(symbol, Tissue, nTPM, TPM) %>%
  dplyr::mutate(logTPM = log(TPM + 0.0001)) %>%
  readr::write_csv(fs::path(tempdir, "tissue_gene_expression.csv")) %>%
  ggplot(aes(symbol, Tissue)) + geom_tile(aes(fill = logTPM),
                                              colour = "white") + scale_fill_gradient(low = "white",
                                                                                      high = "steelblue")+
  labs(x='', y = '') +
  theme_bw() +
  ggtitle(paste("Tissue expression of genes from", filter_context)) +
  guides(fill = guide_legend(title = "Log(nTPM)")) +
  #theme(legend.position="none") +
  theme(plot.title = element_text(hjust = 0.5,size = 20),axis.title = element_text(size=15))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),panel.grid.major= element_blank(),panel.grid.minor = element_blank())

```


```{r kidney-specific-protein-expression}



p <- genes %>%
  dplyr::filter(context == filter_context) %>%
  left_join(normal) %>%
  dplyr::filter(Tissue == filter_tissue) %>%
  dplyr::select(symbol, Tissue, cell_type, value) %>%
  readr::write_csv(fs::path(tempdir, "tissue_specific_protein_hpa.csv")) %>%
  ggplot(aes(symbol, cell_type)) + geom_tile(aes(fill = value),
                                              colour = "white") + scale_fill_gradient(low = "white",
                                                                                      high = "steelblue")+
  labs(x = '', y = '') +
  theme_bw() +
  guides(fill = guide_legend(title = "Protein Level")) +
  ggtitle(paste("Protein detection in", filter_tissue, "for genes from", filter_context)) +
  #theme(legend.position="none") +
  theme(plot.title = element_text(hjust = 0.5,size = 20),axis.title = element_text(size=15)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
  panel.grid.major= element_blank(),panel.grid.minor = element_blank())

tempfile <- fs::path(tempdir, "kidney_structure_protein_detection_renal_new_targets.png")
png(tempfile, width = 960)
print(p)
while (!is.null(dev.list()))  dev.off()
ckanr::resource_create(package_id = tissue_expression_pkg_id,
                     name = "renal_new_targets_kidney structure_protein_detection_heatmap",
                     description = paste("Protein detection by kidney structure for genes",
                                         "in the Renal New Targets 2023 campaign (HPA)"),
                     upload = tempfile,
                     url = ckan_url,
                     key = ckan_key)
unlink(tempfile)

```

## 3.

```{r internally-defined-variants}


```

## 4.



## 5.

```{r gnomad-maf, eval=FALSE}

# fetch MAF from open targets

con <- noctua::dbConnect(noctua::athena(),
                         profile_name = user_profile,
                         s3_staging_dir = s3_staging_dir)

query_string <- paste("select chr_id, position, ref_allele, alt_allele, gene_id_any,",
                      "af.gnomad_afr as African, af.gnomad_amr as American,",
                      "af.gnomad_asj as Ashkenazi, af.gnomad_eas as East_Asian,",
                      "af.gnomad_fin as Finnish, af.gnomad_nfe as Non_Finnish_European,",
                      "af.gnomad_nfe_est as Estonian, af.gnomad_nfe_nwe as NW_European,",
                      "af.gnomad_nfe_onf as Other_Non_Finnish, af.gnomad_nfe_seu as S_European,",
                      "af.gnomad_oth as Other",
                      "from open_targets_genetics.variant_index where gene_id_any in",
                      gene_string)
system.time({res <- noctua::dbExecute(con, query_string)})
maf_output <- noctua::dbFetch(res) %>%
    tibble::as_tibble() %>%
    dplyr::rename(gene_id = gene_id_any,
                  chr = chr_id, pos = position, ref = ref_allele, alt = alt_allele) %>%
    tidyr::unite("label", chr:alt) %>%
    dplyr::select(-gene_id)
noctua::dbClearResult(res)

internal_variants %>%
    left_join(maf_output, by="label")

```


## 6. Disease phenotypes

```{r disease-associations}

# fetch evidence from Open Targets

con <- noctua::dbConnect(noctua::athena(),
                         profile_name = user_profile,
                         s3_staging_dir = s3_staging_dir)

query_string <- paste("select * from",
                      "(select * from open_targets.association_by_overall_direct",
                      "left outer join open_targets.diseases on diseaseid = id)",
                      "where targetid in",
                      gene_string)
system.time({res <- noctua::dbExecute(con, query_string)})
direct_association_output <- noctua::dbFetch(res) %>%
    tibble::as_tibble() %>%
    dplyr::rename(gene_id = targetid) %>%
    dplyr::mutate(association = "direct") %>%
    dplyr::select(gene_id, diseaseid, score, evidencecount, therapeuticareas, name, description, association)
noctua::dbClearResult(res)

query_string <- paste("select * from",
                      "(select * from open_targets.association_by_overall_indirect",
                      "left outer join open_targets.diseases on diseaseid = id)",
                      "where targetid in",
                      gene_string)
system.time({res <- noctua::dbExecute(con, query_string)})
indirect_association_output <- noctua::dbFetch(res) %>%
    tibble::as_tibble() %>%
    dplyr::rename(gene_id = targetid) %>%
    dplyr::mutate(association = "indirect") %>%
    dplyr::select(gene_id, diseaseid, score, evidencecount, therapeuticareas, name, description, association)
noctua::dbClearResult(res)

associations <- direct_association_output %>%
    bind_rows(indirect_association_output) %>%
    group_by(gene_id) %>%
    dplyr::arrange(desc(score))

associations_by_gene <- setNames(associations %>% group_split(),
                                 associations %>% group_keys() %>% pull(gene_id))

lapply(names(associations_by_gene), function(gene) {
    associations_by_gene
})



```



## 7. Mouse models and phenotypes

```{r mouse-models-and-phenotypes-from-file}
knitr::opts_chunk$set(eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE,
                      fig.width=12, fig.height=8)

# FOLLOWING DOES NOT WORK because the columns are not strings. Instead they are
# `JSON-B` formatted structs
# For eventual re-write
# con <- noctua::dbConnect(noctua::athena(),
#                          profile_name = user_profile,
#                          s3_staging_dir = s3_staging_dir)
# 
# 
# query_string <- paste("select * from open_targets.mouse_phenotypes where targetfromsourceid in", gene_string)
# 
# system.time({res <- noctua::dbExecute(con, query_string)})
# 
# output <- noctua::dbFetch(res) %>%
# tibble::as_tibble()
# 
# noctua::dbClearResult(res)

# For the time being, process from local file containing one JSON string per line
# Function to read a JSON file where each line is a JSON string
read_json_lines <- function(filepath) {
  lines <- read_lines(filepath)
  df_list <- map(lines, ~fromJSON(.x, simplifyDataFrame = TRUE))
  lapply(df_list, function(df) {
    df %>%
      map_if(is.data.frame, list) %>%
      as_tibble()
  }) %>% bind_rows()
}

filepath <- fs::path(data_dir, "opentargets/mouse_phenotypes.json")

mouse_models <- read_json_lines(filepath)

# for import into sql table:
# mouse_models <- mouse_models %>% mutate(index = row_number())
# biologicalModels <- mouse_models %>% dplyr::select(index, biological_models) %>% unnest(biologicalModels) 
# modelPhenotypeClasses <- mouse_models %>% unnest(modelPhenotypeClasses)
# modelDetails <- mouse_models %>% select(-biologicalModels, -modelPhenotypeClasses)

mouse_models %>% saveRDS(fs::path(data_dir, "opentargets/mouse_models.rds"))

relevant_models <- mouse_models %>%
  dplyr::filter(targetFromSourceId %in% genelist)

relevant_models %>% saveRDS(fs::path(data_dir, "opentargets/relevant_models.rds"))


model_count <- genes %>%
  dplyr::filter(context == filter_context) %>%
  left_join(relevant_models %>% summarize(number_of_models=n(),
                                          .by = targetFromSourceId),
            by=c("gene_id"="targetFromSourceId")) %>%
  tidyr::replace_na(list("number_of_models" = 0))

model_summary <- setNames(relevant_models %>%
                            group_by(targetFromSourceId) %>%
                            group_split(),
                          relevant_models %>%
                            group_by(targetFromSourceId) %>%
                            group_keys() %>%
                            left_join(genes %>%
                                        dplyr::filter(context == filter_context),
                                      by=c("targetFromSourceId"="gene_id")) %>%
                            pull(symbol))


model_catalog <- lapply(model_summary, function(target) {
  target %>%
    left_join(genes %>%
                dplyr::filter(context == filter_context),
              by=c("targetFromSourceId"="gene_id")) %>%
    dplyr::select(-context, -modelPhenotypeClasses) %>%
    unnest(biologicalModels)
})

```



```{r}
model_catalog[[1]] %>%
  dplyr::select(symbol, targetFromSourceId, targetInModel, targetInModelMgiId,
                targetInModelEnsemblId, modelPhenotypeId, modelPhenotypeLabel,
                id, allelicComposition, geneticBackground) %>%
  flextable::regulartable() %>%
  flextable::theme_box() %>%
  flextable::merge_v(c("symbol", "targetFromSourceId", "targetInModel",
                       "targetInModelMgiId", "targetInModelEnsemblId",
                       "modelPhenotypeId", "modelPhenotypeLabel")) %>%
  flextable::valign(valign = "top", part = "all") %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::set_header_labels(targetFromSourceId = "ENSG",
                               targetInModel = "mouse gene",
                               targetInModelMgiId = "mouse gene MGI",
                               targetInModelEnsemblId = "mouse gene ENS",
                               modelPhenotypeId = "phenotype",
                               id = "model ID",
                               allelicComposition = "allelic composition",
                               geneticBackground = "genetic background"
                               )
```



```{r tissue-specificity, eval=FALSE}
geneset <- targets %>%
  dplyr::filter(context == "Renal New Targets 2023") %>%
  distinct(symbol) %>%
  pull(symbol)

gs <- GeneSet(geneIds = geneset, organism="Homo Sapiens", geneIdType=SymbolIdentifier())
output <- teEnrichment(inputGenes = gs)


seEnrichmentOutput<-output[[1]]
enrichmentOutput<-setNames(data.frame(assay(seEnrichmentOutput),row.names = rowData(seEnrichmentOutput)[,1]), colData(seEnrichmentOutput)[,1])
enrichmentOutput$Tissue<-row.names(enrichmentOutput)
enrichmentOutput %>%
  tibble::as_tibble() %>%
  dplyr::select(Tissue, Tissue.Specific.Genes, fold.change, Log10PValue) %>%
  rename(tissue_specific_genes = Tissue.Specific.Genes,
         fold_change = fold.change)

seExp<-output[[2]][["Kidney"]]
exp<-setNames(data.frame(assay(seExp), row.names = rowData(seExp)[,1]), colData(seExp)[,1])
exp$Gene<-row.names(exp)
exp<-exp %>% gather(key = "Tissue", value = "expression",1:(ncol(exp)-1))

tissue_enrichment <- exp %>%
  tibble::as_tibble() %>%
  pivot_wider(names_from = Tissue, values_from = expression) %>%
  arrange(Gene)

ggplot(exp, aes(Tissue, Gene)) + geom_tile(aes(fill = expression),
                                           colour = "white") + scale_fill_gradient(low = "white",
                                                                                   high = "steelblue")+
  labs(x='', y = '')+
  theme_bw()+
  guides(fill = guide_legend(title = "Log2(TPM)"))+
  #theme(legend.position="none")+
  theme(plot.title = element_text(hjust = 0.5,size = 20),axis.title = element_text(size=15))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),panel.grid.major= element_blank(),panel.grid.minor = element_blank())


```


```{r query-athena}


con <- noctua::dbConnect(noctua::athena(),
                         profile_name = user_profile,
                         s3_staging_dir = s3_staging_dir)


query_string <- "select * from sightline.target_concepts"

system.time({res <- noctua::dbExecute(con, query_string)})

output <- noctua::dbFetch(res)
noctua::dbClearResult(res)

output %>% tibble::as_tibble()
```
---
title: "auto-annotation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{auto-annotation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(AnnotationEngine)
```
